---
title: "Assignment2"
author: "Maxwell Kennady"
date: "October 2, 2019"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(tidycensus)
library(tigris)
library(magrittr)
library(dplyr)
library(leaflet)
library(rgdal)
```


## Texas Residents on Public Health Insurance
Source: U.S. Census Bureau, ACS 5-year, Table B27015 and B01003
Table 1 title: HEALTH INSURANCE COVERAGE STATUS AND TYPE BY HOUSEHOLD INCOME IN THE PAST 12 MONTHS, 
https://data.census.gov/cedsci/table?q=B27015

Table 2 title: TOTAL POPULATION, 
https://data.census.gov/cedsci/table?q=b01003

```{r layer of polygons with variable that changes layer/legend color}

# Reference: https://walkerke.github.io/tidycensus/articles/basic-usage.html
# Reference: https://juliasilge.com/blog/using-tidycensus/

# View Census data tables available
#View(load_variables(2017, "acs5", cache = TRUE))

# persons on public health insurance from census in Texas counties
public.HI <- get_acs(geography = "county", 
                     variables = c("B27015_002"), 
                     state = "TX",
                     geometry = FALSE,
                     year = 2017)

# population of Texas counties
pop <- get_acs(geography = "county", 
                     variables = c("B01003_001"), 
                     state = "TX",
                     geometry = FALSE,
                     year = 2017)

public.HI.pop <- inner_join(public.HI, pop, by = c("GEOID" = "GEOID", "NAME" = "NAME"))

# get percentage of population of county on public health insurance
public.HI.pop %<>%
  mutate(perc_HI = round(estimate.x / estimate.y, digits = 2))


# get tigris data on counties in Texas and merge with health insurance data frame
# Reference: https://github.com/walkerke/tigris
tx.counties <- counties(state = 48)

tx.counties@data <- merge(tx.counties@data, public.HI.pop, sort = FALSE, by = "GEOID", )

# create palettes of colors
# Reference: https://rstudio.github.io/leaflet/colors.html
pal <- colorBin(palette = "Oranges",
                bins = c(0, .05, 0.1, 0.15, 0.2, 0.25, .3, .35, .4),
                domain = tx.counties$perc_HI)

pal2 <- colorBin(palette = "PuRd",
                bins = c(0, .05, 0.1, 0.15, 0.2, 0.25, .3, .35, .4),
                domain = tx.counties$perc_HI)


leaflet(tx.counties) %>%
  addProviderTiles("OpenStreetMap") %>% 
  # Orange shading of public health insurance
  addPolygons(fillColor = ~pal(perc_HI),
              weight = 1,
              opacity = 1,
              color = "black",
              fillOpacity = 0.6,
              popup = ~paste0("<b>", NAMELSAD, "</b> ",
                              "<br>Residents: ", format(estimate.y, big.mark = "," ),
                              "<br>Residents on public health insurance: ", format(estimate.x,big.mark = ","),
                              "<br>Percent of residents on public<br>health insurance: ",
                              perc_HI * 100, "%"),
              group = "Orange") %>%
  # Purple shading of public health insurance
  addPolygons(fillColor = ~pal2(perc_HI),
              weight = 1,
              opacity = 1,
              color = "black",
              fillOpacity = 0.6,
              popup = ~paste0("<b>", NAMELSAD, "</b> ",
                              "<br>Residents: ", format(estimate.y, big.mark = "," ),
                              "<br>Residents on public health insurance: ", format(estimate.x,big.mark = ","),
                              "<br>Percent of residents on public<br>health insurance: ",
                              perc_HI * 100, "%"),
              group = "Purple") %>%
  # Reference: https://rstudio.github.io/leaflet/legends.html
  # Reference: https://stackoverflow.com/questions/38701359/grouped-layer-control-in-leaflet-r
  # Reference: https://rstudio.github.io/leaflet/showhide.html
  addLegend(position = "bottomright",
            pal = pal,
            values = ~perc_HI,
            title = "Percent of county residents<br>on public health insurance",
            labFormat = labelFormat(digits = 0, transform = function(x) 100 * x, suffix = "%" ),
            group = "Orange") %>%
  addLegend(position = "bottomright",
            pal = pal2,
            values = ~perc_HI,
            title = "Percent of county residents<br>on public health insurance",
            labFormat = labelFormat(digits = 0, transform = function(x) 100 * x, suffix = "%" ),
            group = "Purple") %>%
  # allow user to change color of shading and legend
  addLayersControl(overlayGroups = c("Orange", "Purple"),
                   options = layersControlOptions(collapsed=FALSE)) %>%
  # by default, hide the group "Purple"
  hideGroup("Purple")



```





## Natural Gas Pipelines in the United States
```{r layer of lines}

# data source: https://hifld-geoplatform.opendata.arcgis.com/datasets/natural-gas-pipelines?geometry=-109.914%2C26.933%2C-86.249%2C33.570

pipelines <- readOGR("./Natural_Gas_Pipelines/Natural_Gas_Pipelines.shp")

leaflet(pipelines) %>%
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  addPolylines(color = "red",
              weight = 1, 
              group = "Red")

```


## Residents of Texas Counties working in Mining, Quarrying, and Oil and Gas Extraction
Source: U.S. Census Bureau, ACS 5-year, Table C24040
Table:  SEX BY INDUSTRY FOR THE FULL-TIME, YEAR-ROUND CIVILIAN EMPLOYED POPULATION 16 YEARS AND OVER,
https://data.census.gov/cedsci/table?hidePreview=true&q=C24040

```{r layer of points and polygons with functioning layersControl (two basemaps, two layers)}

# data source: https://hifld-geoplatform.opendata.arcgis.com/datasets/oil-refinieries-1
refineries <- read_csv("./Oil_Refineries.csv")

oil <- get_acs(geography = "county", 
                     variables = c("C24040_005"), 
                     state = "TX",
                     geometry = FALSE,
                     year = 2017)

oilFemale <- get_acs(geography = "county", 
                     variables = c("C24040_032"), 
                     state = "TX",
                     geometry = FALSE,
                     year = 2017)

# merge datasets and get percentage of each county's workers in mining/oil industry
oil %<>%
  left_join(oilFemale, by = c("GEOID" = "GEOID", "NAME" = "NAME")) %>%
  left_join(pop, by = c("GEOID" = "GEOID", "NAME" = "NAME")) %>%
  mutate(total_oil = estimate.x + estimate.y, perc_oil = total_oil / estimate)

oil$perc_oil <- as.numeric(round(oil$perc_oil, digits = 2))  * 100

counties.oil <- counties(state = 48)

counties.oil@data <- merge(counties.oil@data, oil, sort = FALSE, by = "GEOID")


pal3 <- colorBin(palette = "Oranges",
                bins = c(0, 2, 4, 6, 8, 12, 14),
                domain = counties.oil$perc_oil)


leaflet(counties.oil) %>%
  addProviderTiles("OpenStreetMap", group = "OSM") %>% 
  addProviderTiles("Stamen.Terrain", group = "Stamen") %>% 
  addPolygons(fillColor = ~pal3(perc_oil),
              weight = 1,
              opacity = 1,
              color = "black",
              fillOpacity = 0.6,
              popup = ~paste0("<b>", NAMELSAD, "</b> ",
                              "<br>Residents working in oil & gas: ", format(total_oil, big.mark = ","),
                              "<br>Percent of residents working in oil & gas: ", perc_oil, "%"),
              group = "Oil/Gas Workers") %>%
  addMarkers(data = refineries, 
             lng = ~LONGITUDE, 
             lat = ~LATITUDE, 
             popup = paste0(refineries$OPERNAME, "<br>", 
                            refineries$ADDRESS, "<br>", 
                            refineries$CITY, ", ", 
                            refineries$STATE, " ", 
                            refineries$ZIP),
             group = "Refineries") %>%
  addLegend(position = "bottomright",
            pal = pal3,
            values = ~perc_oil,
            title = "Percent of residents working in oil & gas",
            labFormat = labelFormat(digits = 0, suffix = "%" )) %>%
  addLayersControl(baseGroups = c("OSM", "Stamen"),
                   overlayGroups = c("Oil/Gas Workers","Refineries"),
                   options = layersControlOptions(collapsed = FALSE))





```




